---
title     : "Exercises: Dependency Management"
subtitle: "Workflows for Reproducible Research with R & Git"
---

```{r child = "content/config/exercises_setup.Rmd"}
```

In this brief set of exercises, you will try some simple solutions for dependency management in `R`. For this, you should continue working with the `R` script you have created and pushed to your *GitHub* repository in the previous sessions.

```{block, box.title = "Exercise 1", box.body = list(fill = "white"), box.icon = "fa-star"}
First, add some basic reproducibility information about the version of `R` and the packages you use to your script. If you want to, you can can exclude information about your `locale`.
```

```{block, opts.label = "clues", box.icon = "fa-lightbulb"}
You can use a simple function from base `R` to get some information about your current session. If you want to exclude information about your `locale`, you need to wrap that function into a call to the `print()` function and provide it with an additional argument.
```

```{r session, solution = TRUE, eval = F, box.title = "Solution 1", solution = TRUE, box.icon = "fa-check"}
print(sessionInfo(), locale = FALSE)
```

```{block, box.title = "Exercise 2", box.body = list(fill = "white"), box.icon = "fa-star"}
Now, use the `groundhog` package to add load the packages you use in your script.
```

```{block, opts.label = "clues", box.icon = "fa-lightbulb"}
Do not forget to specify a date. You can just use today's date.
```

```{r groundhog, solution = TRUE, eval = F, box.title = "Solution 2", solution = TRUE, box.icon = "fa-check"}
library(groundhog)

# adapt the package names to include the ones you use in your script
pkgs <- c("tidyverse", "janitor", "sjPlot")

groundhog.library(pkgs, date = "2023-11-71")
```

If you are done with these exercises very quickly or if they are too easy for you, feel free to change the arguments for the `groundhog.library()` function and see what happens. You can also try to use the `groundhog` package to install a package that is not available on CRAN anymore. Or read some of the detailed documentation on the [`groundhog` package website](https://groundhogr.com/).
