---
title     : "Exercises: Dependency Management"
subtitle: "Workflows for Reproducible Research with R & Git"
---

```{r child = "content/config/exercises_setup.Rmd"}
```

In this brief set of exercises, you will try some simple solutions for dependency management in `R`. For this, you should continue working with the `R` script you have created and pushed to your *GitHub* repository in the previous sessions.

```{block, box.title = "Exercise 1", box.body = list(fill = "white"), box.icon = "fa-star"}
First, add get some information about the version of `R` you are using as well as the names and versions of the `R` packages you have loaded in your current `R` session. Add a comment containing the version of `R` you are using to the `R` script you have created and edited in the previous sessions.
```

```{block, opts.label = "clues", box.icon = "fa-lightbulb"}
You can, of course, get the information about your `R` version and the packages via the GUI of *RStudio* (e.g., in the Console and the *Packages* tab). However, you should should use base `R` functions to get this information.
```

```{r session, solution = TRUE, box.title = "Solution 1", solution = TRUE, box.icon = "fa-check"}
# Only information about your R version
R.version.string

# Detailed information about your current R session
sessionInfo()
```

```{block, box.title = "Exercise 2", box.body = list(fill = "white"), box.icon = "fa-star"}
Now, use the `groundhog` package to load the packages you use in your script.
```

```{block, opts.label = "clues", box.icon = "fa-lightbulb"}
Do not forget to specify a date. You can just use today's date.
```

```{r groundhog, solution = TRUE, eval = F, box.title = "Solution 2", solution = TRUE, box.icon = "fa-check"}
library(groundhog)

# adapt the package names to include the ones you use in your script
pkgs <- c("dplyr", "ggplot2", "scales", "correlation", "sjPlot")

groundhog.library(pkgs, date = "2023-11-17")
```

If you are done with these exercises very quickly or if they are too easy for you, feel free to change some of the arguments for the `groundhog.library()` function and see what happens. You can also try to use the `groundhog` package to install a package that is not available on CRAN anymore. Or read some of the detailed documentation on the [`groundhog` package website](https://groundhogr.com/).
