---
title: "Workflows for Reproducible Research with R & Git"
subtitle: "Recap Day 1"
author: "Johannes Breuer, Bernd Weiss, & Arnim Bleier"
date: "2023-11-17"
presenter: Johannes
editor_options: 
  chunk_output_type: console
---
layout: true

```{r child = "content/config/sessions_setup.Rmd"}
```

---

## What we covered on Day 1

```{r schedule-1, echo = F}
course_content_1
```

---

## What we covered on Day 1

- What is reproducibility? Why does it matter?
- Some computer literacy basics: operating systems, file systems, paths, command line...
- What are `Git` and *GitHub*?
  - Setup
  - Basic usage
- How to use `Git` & *GitHub* with(in) *RStudio*?
- What else can we do with `Git` & *GitHub* besides creating repositories, adding, committing, pushing, and pulling files?

---

class: center, middle

# Any remaining questions or issues from Day 1?

---

## What we will cover today

- Managing dependencies (within `R` and beyond)
- Using *Binder* to share your analyses and results and offer "one-click reproducibility"
- Saving computational environments
- Other topics and tools for reproducible research that we have not covered in this course

---

## What we will cover today

```{r schedule-2, echo = F}
course_content_2
```

---

class: center, middle

# Let's get to it! `r ji("rocket")`